\chapter*{List of Publications}
\pagestyle{plain}
%\chaptermark{ }
%\sectionmark{ }
%\addcontentsline{toc}{chapter}{List of Publications}

\newcounter{pubs_counter}
\begin{list}{[\arabic{pubs_counter}]~}{\usecounter{pubs_counter}}

\item
R.~Bonfigli, A.~Felicetti, E.~Principi, M.~Fagiani, S.~Squartini, and
  F.~Piazza, ``Denoising autoencoders for non-intrusive load monitoring:
  Improvements and comparative evaluation,'' \emph{Energy and Buildings}, to appear.

\item
R.~Bonfigli, E.~Principi, M.~Fagiani, M.~Severini, S.~Squartini, and
  F.~Piazza, ``Non-intrusive load monitoring by using active and reactive power
  in additive factorial hidden markov models,'' \emph{Applied Energy}, vol.
  208, no. Supplement C, pp. 1590 -- 1607, 2017.
  
\item  % NB Insert or not ??
patent, ``Metodo per il monitoraggio non intrusivo del consumo di apparecchiature
elettriche collegate ad una linea di alimentazione comune,'' Domanda numero:
102017000004554, patent pending.

\item
R.~Bonfigli, E.~Principi, S.~Squartini, M.~Fagiani, M.~Severini, and
  F.~Piazza, ``{User-aided Footprint Extraction for Appliance Modelling in
  Non-Intrusive Load Monitoring},'' in \emph{Proc. of the IEEE Symposium Series
  on Computational Intelligence}, Athens, Greece, Dec. 6-9 2016, pp. 1--8.

\item
R.~Bonfigli, M.~Severini, S.~Squartini, M.~Fagiani, and F.~Piazza,
  ``Improving the performance of the {AFAMAP} algorithm for non-intrusive load
  monitoring,'' in \emph{Proc. of the IEEE Congress on Evolutionary Computation
  (CEC)}, Vancouver, Canada, 2016, pp. 303--310.

\item
M.~Fagiani, S.~Squartini, R.~Bonfigli, M.~Severini, and F.~Piazza,
  ``Exploiting temporal features and pressure data for automatic leakage
  detection in smart water grids,'' in \emph{2016 IEEE Congress on Evolutionary
  Computation (CEC)}, July 2016, pp. 295--302.

\item
R.~Bonfigli, S.~Squartini, M.~Fagiani, and F.~Piazza, ``Unsupervised
  algorithms for non-intrusive load monitoring: An up-to-date overview,'' in
  \emph{Proc. of IEEE 15th Int. Conf. on Environment and Electrical Engineering
  (EEEIC)}, June 2015, pp. 1175--1180.

\item
M.~Fagiani, S.~Squartini, R.~Bonfigli, and F.~Piazza, ``Short-term
  load forecasting for smart water and gas grids: A comparative evaluation,''
  in \emph{Environment and Electrical Engineering (EEEIC), 2015 IEEE 15th
  International Conference on}, June 2015, pp. 1198--1203.

\item
G.~Ferroni, R.~Bonfigli, E.~Principi, S.~Squartini, and F.~Piazza, ``A
  deep neural network approach for voice activity detection in multi-room
  domestic scenarios,'' in \emph{2015 International Joint Conference on Neural
  Networks (IJCNN)}, July 2015, pp. 1--8.

\item
E.~Principi, P.~Olivetti, S.~Squartini, R.~Bonfigli, and F.~Piazza,
  ``A floor acoustic sensor for fall classification,'' in \emph{Audio
  Engineering Society Convention 138}, May 2015.

\item
E.~Principi, S.~Squartini, R.~Bonfigli, G.~Ferroni, and F.~Piazza,
  ``An integrated system for voice command recognition and emergency detection
  based on audio signals,'' \emph{Expert Systems with Applications}, vol.~42,
  no.~13, pp. 5668 -- 5683, 2015.

\item
G.~Ferroni, R.~Bonfigli, E.~Principi, S.~Squartini, and F.~Piazza,
  ``Neural networks based methods for voice activity detection in a multi-room
  domestic environment,'' in \emph{XIII AI*IA Symposium on Artificial
  Intelligence}, Dec 2014.
  
\end{list}